---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# GWlasso

<!-- badges: start -->
[![R-CMD-check](https://github.com/nibortolum/GWlasso/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/nibortolum/GWlasso/actions/workflows/R-CMD-check.yaml)
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)
[![CRAN status](https://www.r-pkg.org/badges/version/GWlasso)](https://CRAN.R-project.org/package=GWlasso)
[![Dowloads](https://cranlogs.r-pkg.org/badges/grand-total/GWlasso)](https://CRAN.R-project.org/package=GWlasso)
<!-- badges: end -->

The goal of GWlasso is to provides a set of functions to perform Geographically weighted lasso. It was originally thought to be used in palaeoecological settings but can be used to other extents. 


## Installation

You can install the development version of GWlasso from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("nibortolum/GWlasso")
```

You can install the stable version directly from CRAN with

``` r
install.packages("GWlasso")

```


## Example

This is a basic example on how to run a GWlasso pipeline:

```{r example, include=TRUE, eval = FALSE}
library(GWlasso)

## compute a distance matrix from a set of coordinates
distance_matrix <- compute_distance_matrix <- function(coords, method = "euclidean", add.noise = FALSE)

## compute the optimal bandwidth 
  myst.est <- gwl_bw_estimation(x.var = predictors_df, 
                              y.var = y_vector,
                              dist.mat = distance_matrix,
                              adaptive = TRUE,
                              adptbwd.thresh = 0.1,
                              kernel = "bisquare",
                              alpha = 1,
                              progress = TRUE,
                              n=40,
                              nfolds = 5)

## Compute the optimal model
my.gwl.fit <- gwl_fit(myst.est$bw,
                      x.var = data.sample[,-1], 
                      y.var = data.sample$WTD,
                      kernel = "bisquare",
                      dist.mat = distance_matrix, 
                      alpha = 1, 
                      adaptive = TRUE, progress = T)

## make predictions 

predicted_values <- predict(my.gwl.fit, newdata = new_data, newcoords = new_coords)
```



